### 2015-12-06 Reference.tex

%boundary conditions imposing for Hyperbolic system of equaiton.
@article{thompson1990time,
  title={Time-dependent boundary conditions for hyperbolic systems, II},
  author={Thompson, Kevin W},
  journal={Journal of Computational Physics},
  volume={89},
  number={2},
  pages={439--461},
  year={1990},
  publisher={Elsevier}
}
@book{godlewski2013numerical,
  title={Numerical approximation of hyperbolic systems of conservation laws},
  author={Godlewski, Edwige and Raviart, Pierre-Arnaud},
  volume={118},
  year={2013},
  publisher={Springer Science \& Business Media}
}
@article{kreiss1970initial,
  title={Initial boundary value problems for hyperbolic systems},
  author={Kreiss, Heinz-Otto},
  journal={Communications on Pure and Applied Mathematics},
  volume={23},
  number={3},
  pages={277--298},
  year={1970},
  publisher={Wiley Online Library}
}
@article{majda1975initial,
  title={Initial-boundary value problems for hyperbolic equations with uniformly characteristic boundary},
  author={Majda, Andrew and Osher, Stanley},
  journal={Communications on Pure and Applied Mathematics},
  volume={28},
  number={5},
  pages={607--675},
  year={1975},
  publisher={Wiley Online Library}
}
@article{ralston1971note,
  title={Note on a paper of Kreiss},
  author={Ralston, James V},
  journal={Communications on Pure and Applied Mathematics},
  volume={24},
  number={6},
  pages={759--762},
  year={1971},
  publisher={Wiley Online Library}
}
@article{benabdallah1987problemes,
  title={Probl{\`e}mes aux limites pour des syst{\`e}mes hyperboliques non lin{\'e}aires de deux {\'e}quations {\`a} une dimension d'espace},
  author={Benabdallah, Assia and Serre, Denis and LIONS, J-L},
  journal={Comptes rendus de l'Acad{\'e}mie des sciences. S{\'e}rie 1, Math{\'e}matique},
  volume={305},
  number={15},
  pages={677--680},
  year={1987},
  publisher={Elsevier}
}
@book{dubois1989boundary,
  title={Boundary conditions for nonlinear hyperbolic systems of conservation laws},
  author={Dubois, Fran{\c{c}}ois and Le Floch, Philippe},
  year={1989},
  publisher={Springer}
}
@techreport{oliger1976theoretical,
  title={Theoretical and Practical Aspects of Some Initial-Boundary Value Problems in Fluid Dynamics.},
  author={Oliger, Joseph and Sundstrom, Arne},
  year={1976},
  institution={DTIC Document}
}
%The paper that I started from to get a basic idea on  IBCP.
@article{guaily2013boundary,
  title={Boundary conditions for hyperbolic systems of partial differentials equations},
  author={Guaily, Amr G and Epstein, Marcelo},
  journal={Journal of advanced research},
  volume={4},
  number={4},
  pages={321--329},
  year={2013},
  publisher={Elsevier}
}
%Books on CFD
@book{pletcher2012computational,
  title={Computational fluid mechanics and heat transfer},
  author={Pletcher, Richard H and Tannehill, John C and Anderson, Dale},
  year={2012},
  publisher={CRC Press}
}

%Very old paper on volcano plume---> useful as in review.
@inproceedings{morton1956turbulent,
  title={Turbulent gravitational convection from maintained and instantaneous sources},
  author={Morton, BR and Taylor, Geoffrey and Turner, JS},
  booktitle={Proceedings of the Royal Society of London A: Mathematical, Physical and Engineering Sciences},
  volume={234},
  number={1196},
  pages={1--23},
  year={1956},
  organization={The Royal Society}
}
@book{bullard1962volcanoes,
  title={Volcanoes: in history, in theory, in eruption},
  author={Bullard, Fred Mason},
  year={1962},
  publisher={Wiley Online Library}
}
@article{settle1978volcanic,
  title={Volcanic eruption clouds and the thermal power output of explosive eruptions},
  author={Settle, Mark},
  journal={Journal of Volcanology and Geothermal Research},
  volume={3},
  number={3},
  pages={309--324},
  year={1978},
  publisher={Elsevier}
}
@article{wilson1978control,
  title={The control of volcanic column heights by eruption energetics and dynamics},
  author={Wilson, L and Sparks, RSJ and Huang, TC and Watkins, ND},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={83},
  number={B4},
  pages={1829--1836},
  year={1978},
  publisher={Wiley Online Library}
}
@article{woods1988fluid,
  title={The fluid dynamics and thermodynamics of eruption columns},
  author={Woods, AW},
  journal={Bulletin of Volcanology},
  volume={50},
  number={3},
  pages={169--193},
  year={1988},
  publisher={Springer}
}
@article{woods1995decompression,
  title={The decompression of volcanic jets in a crater during explosive volcanic eruptions},
  author={Woods, Andrew W and Bower, Sally M},
  journal={Earth and Planetary Science Letters},
  volume={131},
  number={3},
  pages={189--205},
  year={1995},
  publisher={Elsevier}
}
@book{sparks1997volcanic,
  title={Volcanic plumes},
  author={Sparks, Robert Stephen John and Bursik, MI and Carey, SN and Gilbert, Jennifer and Glaze, LS and Sigurdsson, H and Woods, AW},
  year={1997},
  publisher={Wiley}
}
@article{woods1993moist,
  title={Moist convection and the injection of volcanic ash into the atmosphere},
  author={Woods, Andrew W},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={98},
  number={B10},
  pages={17627--17636},
  year={1993},
  publisher={Wiley Online Library}
}
@article{koyaguchi1996formation,
  title={On the formation of eruption columns following explosive mixing of magma and surface-water},
  author={Koyaguchi, Takehiro and Woods, Andrew W},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={101},
  number={B3},
  pages={5561--5574},
  year={1996},
  publisher={Wiley Online Library}
}
@article{woods1991particle,
  title={Particle fallout, thermal disequilibrium and volcanic plumes},
  author={Woods, Andrew W and Bursik, Marcus I},
  journal={Bulletin of Volcanology},
  volume={53},
  number={7},
  pages={559--570},
  year={1991},
  publisher={Springer}
}

@article{valentine1989numerical,
  title={Numerical models of Plinian eruption columns and pyroclastic flows},
  author={Valentine, Greg A and Wohletz, Kenneth H},
  journal={J. Geophys. Res},
  volume={94},
  pages={1867--1887},
  year={1989}
}
%papers on Volcano plume simulation
@article{pelantinumerical,
  title={Numerical Simulation of Volcanic Jets},
  author={Pelanti, Marica and LeVeque, Randall J}
}
@article{bursik2001effect,
  title={Effect of wind on the rise height of volcanic plumes},
  author={Bursik, M},
  journal={Geophys. Res. Lett},
  volume={28},
  number={18},
  pages={3621--3624},
  year={2001}
}
@article{mastin2007user,
  title={A user-friendly one-dimensional model for wet volcanic plumes},
  author={Mastin, Larry G},
  journal={Geochemistry, Geophysics, Geosystems},
  volume={8},
  number={3},
  year={2007},
  publisher={Wiley Online Library}
}
@book{lockwood2013volcanoes,
  title={Volcanoes: global perspectives},
  author={Lockwood, John P and Hazlett, Richard W},
  year={2013},
  publisher={John Wiley \& Sons}
}
@article{wohletz1984hydrodynamic,
  title={Hydrodynamic aspects of caldera-forming eruptions: Numerical models},
  author={Wohletz, KH and McGetchin, TR and Sandford, MT and Jones, EM},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={89},
  number={B10},
  pages={8269--8285},
  year={1984},
  publisher={Wiley Online Library}
}
@article{valentine1989numerical,
  title={Numerical models of Plinian eruption columns and pyroclastic flows},
  author={Valentine, Greg A and Wohletz, Kenneth H},
  journal={J. Geophys. Res},
  volume={94},
  pages={1867--1887},
  year={1989}
}
@article{dobran1993numerical,
  title={Numerical simulation of collapsing volcanic columns},
  author={Dobran, Flavio and Neri, Augusto and Macedonio, Giovanni},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={98},
  number={B3},
  pages={4231--4259},
  year={1993},
  publisher={Wiley Online Library}
}
@article{neri1994influence,
  title={Influence of eruption parameters on the thermofluid dynamics of collapsing volcanic columns},
  author={Neri, Augusto and Dobran, Flavio},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={99},
  number={B6},
  pages={11833--11857},
  year={1994},
  publisher={Wiley Online Library}
}
@article{neri1996numerical,
  title={Numerical simulation of collapsing volcanic columns with particles of two sizes},
  author={Neri, Augusto and Macedonio, Giovanni},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={101},
  number={B4},
  pages={8153--8174},
  year={1996},
  publisher={Wiley Online Library}
}
@article{oberhuber1998volcanic,
  title={Volcanic plume simulation on large scales},
  author={Oberhuber, Josef M and Herzog, Michael and Graf, Hans-F and Schwanke, Karsten},
  journal={Journal of Volcanology and Geothermal Research},
  volume={87},
  number={1},
  pages={29--53},
  year={1998},
  publisher={Elsevier}
}
@article{clarke2002transient,
  title={Transient dynamics of vulcanian explosions and column collapse},
  author={Clarke, AB and Voight, B and Neri, A and Macedonio, G},
  journal={Nature},
  volume={415},
  number={6874},
  pages={897--901},
  year={2002},
  publisher={Nature Publishing Group}
}
@article{neri1998role,
  title={The role of magma composition and water content in explosive eruptions: 2. Pyroclastic dispersion dynamics},
  author={Neri, Augusto and Papale, Paolo and Macedonio, Giovanni},
  journal={Journal of volcanology and geothermal research},
  volume={87},
  number={1},
  pages={95--115},
  year={1998},
  publisher={Elsevier}
}
@article{neri2002mass,
  title={Mass partition during collapsing and transitional columns by using numerical simulations},
  author={Neri, Augusto and Di Muro, Andrea and Rosi, Mauro},
  journal={Journal of volcanology and geothermal research},
  volume={115},
  number={1},
  pages={1--18},
  year={2002},
  publisher={Elsevier}
}
@article{neri2003coupled,
  title={Coupled conduit and atmospheric dispersal dynamics of the AD 79 Plinian eruption of Vesuvius},
  author={Neri, Augusto and Papale, Paolo and Del Seppia, Dario and Santacroce, Roberto},
  journal={Journal of volcanology and geothermal research},
  volume={120},
  number={1},
  pages={141--160},
  year={2003},
  publisher={Elsevier}
}
@article{neri2003multiparticle,
  title={Multiparticle simulation of collapsing volcanic columns and pyroclastic flow},
  author={Neri, Augusto and Esposti Ongaro, Tomaso and Macedonio, Giovanni and Gidaspow, Dimitri},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={108},
  number={B4},
  year={2003},
  publisher={Wiley Online Library}
}
%PADC
@article{neri20074d,
  title={4D simulation of explosive eruption dynamics at Vesuvius},
  author={Neri, Augusto and Esposti Ongaro, Tomaso and Menconi, Gianluca and De'Michieli Vitturi, Mattia and Cavazzoni, Carlo and Erbacci, Giovanni and Baxter, Peter J},
  journal={Geophysical research letters},
  volume={34},
  number={4},
  year={2007},
  publisher={Wiley Online Library}
}
%I need add most recent paper on ATHAM, PADC, and Japanese Model.
%Japanese Model
@article{suzuki2005numerical,
  title={A numerical study of turbulent mixing in eruption clouds using a three-dimensional fluid dynamics model},
  author={Suzuki, Yujiro J and Koyaguchi, Takehiro and Ogawa, Masaki and Hachisu, Izumi},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={110},
  number={B8},
  year={2005},
  publisher={Wiley Online Library}
}

@book{fischer2013mixing,
  title={Mixing in inland and coastal waters},
  author={Fischer, Hugo B and List, John E and Koh, C Robert and Imberger, Jorg and Brooks, Norman H},
  year={2013},
  publisher={Elsevier}
}

@article{suzuki2010numerical,
  title={Numerical determination of the efficiency of entrainment in volcanic eruption columns},
  author={Suzuki, YJ and Koyaguchi, T},
  journal={Geophysical Research Letters},
  volume={37},
  number={5},
  year={2010},
  publisher={Wiley Online Library}
}
@article{suzuki20133d,
  title={3D numerical simulation of volcanic eruption clouds during the 2011 Shinmoe-dake eruptions},
  author={Suzuki, Yujiro J and Koyaguchi, Takehiro},
  journal={Earth, Planets and Space},
  volume={65},
  number={6},
  pages={581--589},
  year={2013},
  publisher={Springer}
}
@article{suzuki2009three,
  title={A three-dimensional numerical simulation of spreading umbrella clouds},
  author={Suzuki, YJ and Koyaguchi, T},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={114},
  number={B3},
  year={2009},
  publisher={Wiley Online Library}
}
@article{koyaguchi2010effects,
  title={Effects of the crater on eruption column dynamics},
  author={Koyaguchi, Takehiro and Suzuki, Yujiro J and Kozono, Tomofumi},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={115},
  number={B7},
  year={2010},
  publisher={Wiley Online Library}
}
@article{koyaguchi2009effect,
  title={The effect of intensity of turbulence in umbrella cloud on tephra dispersion during explosive volcanic eruptions: Experimental and numerical approaches},
  author={Koyaguchi, Takehiro and Ochiai, Kiyokatsu and Suzuki, Yujiro J},
  journal={Journal of Volcanology and Geothermal Research},
  volume={186},
  number={1},
  pages={68--78},
  year={2009},
  publisher={Elsevier}
}
@article{suzuki20123,
  title={3-D numerical simulations of eruption column collapse: Effects of vent size on pressure-balanced jet/plumes},
  author={Suzuki, Yujiro J and Koyaguchi, Takehiro},
  journal={Journal of Volcanology and Geothermal Research},
  volume={221},
  pages={1--13},
  year={2012},
  publisher={Elsevier}
}

%ATHAM 
@article{oberhuber1998volcanic,
  title={Volcanic plume simulation on large scales},
  author={Oberhuber, Josef M and Herzog, Michael and Graf, Hans-F and Schwanke, Karsten},
  journal={Journal of Volcanology and Geothermal Research},
  volume={87},
  number={1},
  pages={29--53},
  year={1998},
  publisher={Elsevier}
}
@article{herzog1998effect,
  title={The effect of phase changes of water on the development of volcanic plumes},
  author={Herzog, Michael and Graf, Hans-F and Textor, Christiane and Oberhuber, Josef M},
  journal={Journal of Volcanology and Geothermal Research},
  volume={87},
  number={1},
  pages={55--74},
  year={1998},
  publisher={Elsevier}
}
@article{langmann1997radiative,
  title={Radiative forcing of sulfate aerosols as determined by a regional circulation-chemistry transport model},
  author={Langmann, B{\"a}rbel and Herzog, Michael and Graf, Hans-F},
  journal={Journal of Aerosol Science},
  volume={28},
  pages={S389--S390},
  year={1997},
  publisher={Pergamon}
}
@article{trentmann2002simulation,
  title={Simulation of a biomass-burning plume: Comparison of model results with observations},
  author={Trentmann, J and Andreae, MO and Graf, H-F and Hobbs, PV and Ottmar, RD and Trautmann, T},
  journal={Journal of Geophysical Research: Atmospheres (1984--2012)},
  volume={107},
  number={D2},
  pages={AAC--5},
  year={2002},
  publisher={Wiley Online Library}
}
@article{textor2003injection,
  title={Injection of gases into the stratosphere by explosive volcanic eruptions},
  author={Textor, Christiane and Graf, Hans-F and Herzog, Michael and Oberhuber, JM},
  journal={Journal of Geophysical Research: Atmospheres (1984--2012)},
  volume={108},
  number={D19},
  year={2003},
  publisher={Wiley Online Library}
}
@article{herzog2003prognostic,
  title={A prognostic turbulence scheme for the nonhydrostatic plume model ATHAM},
  author={Herzog, Michael and Oberhuber, Josef M and Graf, Hans-F},
  journal={Journal of the atmospheric sciences},
  volume={60},
  number={22},
  pages={2783--2796},
  year={2003}
}

@article{textor2006volcanic1,
  title={Volcanic particle aggregation in explosive eruption columns. Part I: Parameterization of the microphysics of hydrometeors and ash},
  author={Textor, Christiane and Graf, Hans-F and Herzog, Michael and Oberhuber, Josef M and Rose, William I and Ernst, Gerald GJ},
  journal={Journal of Volcanology and Geothermal Research},
  volume={150},
  number={4},
  pages={359--377},
  year={2006},
  publisher={Elsevier}
}
@article{textor2006volcanic2,
  title={Volcanic particle aggregation in explosive eruption columns. Part II: Numerical experiments},
  author={Textor, C and Graf, Hans-F and Herzog, Michael and Oberhuber, Josef M and Rose, William I and Ernst, GGJ},
  journal={Journal of Volcanology and Geothermal Research},
  volume={150},
  number={4},
  pages={378--394},
  year={2006},
  publisher={Elsevier}
}
@book{kobs2009modeling,
  title={Modeling particle motion and near-vent deposition in explosive volcanic eruptions},
  author={Kobs, Shannon},
  year={2009},
  publisher={State University of New York at Buffalo}
}
@article{herzog2010applying,
  title={Applying the three-dimensional model ATHAM to volcanic plumes: Dynamic of large co-ignimbrite eruptions and associated injection heights for volcanic gases},
  author={Herzog, Michael and Graf, Hans-F},
  journal={Geophysical Research Letters},
  volume={37},
  number={19},
  year={2010},
  publisher={Wiley Online Library}
}
@article{bursik2009volcanic,
  title={Volcanic plumes and wind: Jetstream interaction examples and implications for air traffic},
  author={Bursik, MI and Kobs, SE and Burns, A and Braitseva, OA and Bazanova, LI and Melekestsev, IV and Kurbatov, A and Pieri, DC},
  journal={Journal of Volcanology and Geothermal Research},
  volume={186},
  number={1},
  pages={60--67},
  year={2009},
  publisher={Elsevier}
}
@article{van2015hail,
  title={Hail formation triggers rapid ash aggregation in volcanic plumes},
  author={Van Eaton, Alexa R and Mastin, Larry G and Herzog, Michael and Schwaiger, Hans F and Schneider, David J and Wallace, Kristi L and Clarke, Amanda B},
  journal={Nature communications},
  volume={6},
  year={2015},
  publisher={Nature Publishing Group}
}

%review type papers on Volcano
@article{bonadonna2012future,
  title={Future developments in modelling and monitoring of volcanic ash clouds: outcomes from the first IAVCEI-WMO workshop on Ash Dispersal Forecast and Civil Aviation},
  author={Bonadonna, Costanza and Folch, Arnau and Loughlin, Susan and Puempel, Herbert},
  journal={Bulletin of volcanology},
  volume={74},
  number={1},
  pages={1--10},
  year={2012},
  publisher={Springer}
}

%exmperimental or observation type of paper on volcano plume
@article{chojnicki2014flow,
  title={The flow structure of jets from transient sources and implications for modeling short-duration explosive volcanic eruptions},
  author={Chojnicki, KN and Clarke, AB and Adrian, RJ and Phillips, JC},
  journal={Geochemistry, Geophysics, Geosystems},
  volume={15},
  number={12},
  pages={4831--4845},
  year={2014},
  publisher={Wiley Online Library}
}

%1D plume softwares
@misc { 114,
	title = {puffin},
	month = {Jan},
	url = {https://vhub.org/resources/114},
	year = {2013},
	author = {Bursik , Marcus I and Carbonara , Alejandro Uriel and Zawicki , Sean Michael}
}
@misc { 267,
	title = {bent: A model of plumes in crossflow},
	month = {Nov},
	url = {https://vhub.org/resources/267},
	year = {2010},
	author = {Bursik , Marcus I}
}
@misc { 1194,
	title = {Plumeria 2.3.1},
	month = {Dec},
	url = {https://vhub.org/resources/1194},
	year = {2011},
	author = {Mastin , Larry G}
}
@misc { 3541,
	title = {PlumeMoM},
	month = {Mar},
	url = {https://vhub.org/resources/3541},
	year = {2015},
	author = {Vitturi , Mattia de' Michieli}
}
@misc{draxler2015hysplit,
  title={HYSPLIT (HYbrid Single-Particle Lagrangian Integrated Trajectory) Model access via NOAA ARL READY Website (). NOAA Air Resources Laboratory, Silver Spring},
  author={Draxler, R.R. and Rolph, G.D.},
  year={2015},
  publisher={MD}
}
@misc{rolph2015ready,
  title={Real-time Environmental Applications and Display sYstem (READY) Website (http://ready.arl.noaa.gov). NOAA Air Resources Laboratory, Silver Spring,},
  author={Rolph, G.D.},
  year={2015},
  publisher={MD}
}

%SPH
@article{gingold1977smoothed,
  title={Smoothed particle hydrodynamics: theory and application to non-spherical stars},
  author={Gingold, Robert A and Monaghan, Joseph J},
  journal={Monthly notices of the royal astronomical society},
  volume={181},
  number={3},
  pages={375--389},
  year={1977},
  publisher={Oxford University Press}
}
@article{lucy1977numerical,
  title={A numerical approach to the testing of the fission hypothesis},
  author={Lucy, Leon B},
  journal={The astronomical journal},
  volume={82},
  pages={1013--1024},
  year={1977}
}
@article{monaghan1992smoothed,
  title={Smoothed particle hydrodynamics},
  author={Monaghan, Joe J},
  journal={Annual review of astronomy and astrophysics},
  volume={30},
  pages={543--574},
  year={1992}
}
@article{cleary1999conduction,
  title={Conduction modelling using smoothed particle hydrodynamics},
  author={Cleary, Paul W and Monaghan, Joseph J},
  journal={Journal of Computational Physics},
  volume={148},
  number={1},
  pages={227--264},
  year={1999},
  publisher={Elsevier}
}
@article{monaghan2005smoothed,
  title={Smoothed particle hydrodynamics},
  author={Monaghan, Joe J},
  journal={Reports on progress in physics},
  volume={68},
  number={8},
  pages={1703},
  year={2005},
  publisher={IOP Publishing}
}
@article{rosswog2009astrophysical,
  title={Astrophysical smooth particle hydrodynamics},
  author={Rosswog, Stephan},
  journal={New Astronomy Reviews},
  volume={53},
  number={4},
  pages={78--104},
  year={2009},
  publisher={Elsevier}
}
@article{ellero2012smoothed,
  title={Smoothed Particle Hydrodynamics},
  author={Ellero, M},
  journal={Fluid-Structure Interactions in Soft-Matter Systems: From the Mesoscale to the Macroscale},
  pages={22},
  year={2012}
}
@article{monaghan2012smoothed,
  title={Smoothed particle hydrodynamics and its diverse applications},
  author={Monaghan, JJ},
  journal={Annual Review of Fluid Mechanics},
  volume={44},
  pages={323--346},
  year={2012},
  publisher={Annual Reviews}
}
@article{saitoh2013density,
  title={A density-independent formulation of smoothed particle hydrodynamics},
  author={Saitoh, Takayuki R and Makino, Junichiro},
  journal={The Astrophysical Journal},
  volume={768},
  number={1},
  pages={44},
  year={2013},
  publisher={IOP Publishing}
}
@techreport{takayuki2012density,
  title={A Density Independent Formulation of Smoothed Particle Hydrodynamics},
  author={Takayuki, Saitoh R and Makino, Junichiro},
  year={2012}
}
@article{price2012smoothed,
  title={Smoothed particle hydrodynamics and magnetohydrodynamics},
  author={Price, Daniel J},
  journal={Journal of Computational Physics},
  volume={231},
  number={3},
  pages={759--794},
  year={2012},
  publisher={Elsevier}
}
@article{monaghan1985extrapolating,
  title={Extrapolating B splines for interpolation},
  author={Monaghan, JJ},
  journal={Journal of Computational Physics},
  volume={60},
  number={2},
  pages={253--262},
  year={1985},
  publisher={Elsevier}
}
@article{monaghan1995sph,
  title={SPH simulation of multi-phase flow},
  author={Monaghan, JJ and Kocharyan, A},
  journal={Computer Physics Communications},
  volume={87},
  number={1},
  pages={225--235},
  year={1995},
  publisher={Elsevier}
}
@article{morris1997modeling,
  title={Modeling low Reynolds number incompressible flows using SPH},
  author={Morris, Joseph P and Fox, Patrick J and Zhu, Yi},
  journal={Journal of computational physics},
  volume={136},
  number={1},
  pages={214--226},
  year={1997},
  publisher={Elsevier}
}

@article{chen1999improvement,
  title={An improvement for tensile instability in smoothed particle hydrodynamics},
  author={Chen, JK and Beraun, JE and Jih, CJ},
  journal={Computational Mechanics},
  volume={23},
  number={4},
  pages={279--287},
  year={1999},
  publisher={Springer}
}
@article{adami2012generalized,
  title={A generalized wall boundary condition for smoothed particle hydrodynamics},
  author={Adami, S and Hu, XY and Adams, NA},
  journal={Journal of Computational Physics},
  volume={231},
  number={21},
  pages={7057--7075},
  year={2012},
  publisher={Elsevier}
}
@article{kumar2013parallel,
  title={Parallel Godunov smoothed particle hydrodynamics (SPH) with improved treatment of Boundary Conditions and an application to granular flows},
  author={Kumar, Dinesh and Patra, Abani K and Pitman, E Bruce and Chi, H},
  journal={Computer Physics Communications},
  volume={184},
  number={10},
  pages={2277--2286},
  year={2013},
  publisher={Elsevier}
}
@article{monaghan2009sph,
  title={SPH particle boundary forces for arbitrary boundaries},
  author={Monaghan, Joe J and Kajtar, Jules B},
  journal={Computer Physics Communications},
  volume={180},
  number={10},
  pages={1811--1820},
  year={2009},
  publisher={Elsevier}
}
@article{yildiz2009sph,
  title={SPH with the multiple boundary tangent method},
  author={Yildiz, M and Rook, RA and Suleman, A},
  journal={International journal for numerical methods in engineering},
  volume={77},
  number={10},
  pages={1416--1438},
  year={2009},
  publisher={Wiley Online Library}
}
@inproceedings{ihmsen2011parallel,
  title={A Parallel SPH Implementation on Multi-Core CPUs},
  author={Ihmsen, Markus and Akinci, Nadir and Becker, Markus and Teschner, Matthias},
  booktitle={Computer Graphics Forum},
  volume={30},
  number={1},
  pages={99--112},
  year={2011},
  organization={Wiley Online Library}
}
@article{herault2010sph,
  title={Sph on gpu with cuda},
  author={H{\'e}rault, Alexis and Bilotta, Giuseppe and Dalrymple, Robert A},
  journal={Journal of Hydraulic Research},
  volume={48},
  number={S1},
  pages={74--79},
  year={2010},
  publisher={Taylor \& Francis}
}
@article{crespo2011gpus,
  title={GPUs, a new tool of acceleration in CFD: efficiency and reliability on smoothed particle hydrodynamics methods},
  author={Crespo, AJC and Dominguez, Jose M and Barreiro, Anxo and G{\'o}mez-Gesteira, Moncho and Rogers, Benedict D},
  journal={PLoS One},
  volume={6},
  number={6},
  pages={e20685},
  year={2011},
  publisher={Public Library of Science}
}
@article{dominguez2013optimization,
  title={Optimization strategies for CPU and GPU implementations of a smoothed particle hydrodynamics method},
  author={Dom{\'\i}nguez, Jose M and Crespo, Alejandro JC and G{\'o}mez-Gesteira, Moncho},
  journal={Computer Physics Communications},
  volume={184},
  number={3},
  pages={617--627},
  year={2013},
  publisher={Elsevier}
}
@article{loren2014two,
  title={Two-fluid dust and gas mixtures in smoothed particle hydrodynamics: a semi-implicit approach},
  author={Lor{\'e}n-Aguilar, Pablo and Bate, Matthew R},
  journal={Monthly Notices of the Royal Astronomical Society},
  volume={443},
  number={1},
  pages={927--945},
  year={2014},
  publisher={Oxford University Press}
}
@article{rook2007modeling,
  title={Modeling transient heat transfer using SPH and implicit time integration},
  author={Rook, Rusty and Yildiz, Mehmet and Dost, Sadik},
  journal={Numerical Heat Transfer, Part B: Fundamentals},
  volume={51},
  number={1},
  pages={1--23},
  year={2007},
  publisher={Taylor \& Francis}
}

@article{vacondio2012accurate,
  title={Accurate particle splitting for smoothed particle hydrodynamics in shallow water with shock capturing},
  author={Vacondio, R and Rogers, BD and Stansby, PK},
  journal={International Journal for Numerical Methods in Fluids},
  volume={69},
  number={8},
  pages={1377--1410},
  year={2012},
  publisher={Wiley Online Library}
}
@article{feldman2007dynamic,
  title={Dynamic refinement and boundary contact forces in SPH with applications in fluid flow problems},
  author={Feldman, J and Bonet, J},
  journal={International Journal for Numerical Methods in Engineering},
  volume={72},
  number={3},
  pages={295--324},
  year={2007},
  publisher={Wiley Online Library}
}
@article{lastiwka2009permeable,
  title={Permeable and non-reflecting boundary conditions in SPH},
  author={Lastiwka, Martin and Basa, Mihai and Quinlan, Nathan J},
  year={2009},
  publisher={Wiley}
}
@article{colagrossi2005meshless,
  title={A meshless Lagrangian method for free-surface and interface flows with fragmentation},
  author={Colagrossi, Andrea},
  year={2005},
  publisher={La Sapienza}
}
@article{molteni2009simple,
  title={A simple procedure to improve the pressure evaluation in hydrodynamic context using the SPH},
  author={Molteni, Diego and Colagrossi, Andrea},
  journal={Computer Physics Communications},
  volume={180},
  number={6},
  pages={861--872},
  year={2009},
  publisher={Elsevier}
}
@article{kondo2011improvement,
  title={Improvement of stability in moving particle semi-implicit method},
  author={Kondo, Masahiro and Koshizuka, Seiichi},
  journal={International Journal for Numerical Methods in Fluids},
  volume={65},
  number={6},
  pages={638--654},
  year={2011},
  publisher={Wiley Online Library}
}
@article{macia2011theoretical,
  title={Theoretical analysis of the no-slip boundary condition enforcement in SPH methods},
  author={Maci{\'a}, Fabricio and Antuono, Matteo and Gonz{\'a}lez, Leo M and Colagrossi, Andrea},
  journal={Progress of theoretical physics},
  volume={125},
  number={6},
  pages={1091--1121},
  year={2011},
  publisher={Oxford University Press}
}
@article{gotoh2004sph,
  title={SPH-LES model for numerical investigation of wave interaction with partially immersed breakwater},
  author={Gotoh, Hitoshi and Shao, Songdong and Memita, Tetsu},
  journal={Coastal Engineering Journal},
  volume={46},
  number={01},
  pages={39--63},
  year={2004},
  publisher={World Scientific}
}
@article{shao2006sph,
  title={SPH computation of plunging waves using a 2-D sub-particle scale (SPS) turbulence model},
  author={Shao, Songdong and Ji, Changming},
  journal={International Journal for numerical methods in fluids},
  volume={51},
  number={8},
  pages={913--936},
  year={2006},
  publisher={Wiley Online Library}
}
@article{hu2009constant,
  title={A constant-density approach for incompressible multi-phase SPH},
  author={Hu, XY and Adams, NA},
  journal={Journal of Computational Physics},
  volume={228},
  number={6},
  pages={2082--2091},
  year={2009},
  publisher={Elsevier}
}

@article{tong2014incompressible,
  title={An incompressible multi-phase smoothed particle hydrodynamics (SPH) method for modelling thermocapillary flow},
  author={Tong, Mingming and Browne, David J},
  journal={International Journal of Heat and Mass Transfer},
  volume={73},
  pages={284--292},
  year={2014},
  publisher={Elsevier}
}
%GSPH
@article{inutsuka2002reformulation,
  title={Reformulation of smoothed particle hydrodynamics with Riemann solver},
  author={Inutsuka, Shu-ichiro},
  journal={Journal of Computational Physics},
  volume={179},
  number={1},
  pages={238--267},
  year={2002},
  publisher={Elsevier}
}
@inproceedings{borgani2012hydrodynamic,
  title={Hydrodynamic Simulations with the Godunov SPH},
  author={Borgani, S and Murante, G and Brunino, R and Cha, S-H},
  booktitle={Advances in Computational Astrophysics: Methods, Tools, and Outcome},
  volume={453},
  pages={259},
  year={2012}
}

%%Reformulate SPH by Galerkin
@article{diltsg1999moving,
  title={Moving-least-squares-particle hydrodynamics—I Consistency and stability},
  author={Dilts, Gary A},
  journal={International Journal for Numerical Methods in Engineering},
  volume={44},
  number={8},
  pages={1115--1155},
  year={1999},
  publisher={Wiley Online Library}
}

@article{monaghan1985refined,
  title={A refined particle method for astrophysical problems},
  author={Monaghan, Joseph J and Lattanzio, John C},
  journal={Astronomy and astrophysics},
  volume={149},
  pages={135--143},
  year={1985}
}

%%turbulent && SPH
@article{shao2006sph,
  title={SPH computation of plunging waves using a 2-D sub-particle scale (SPS) turbulence model},
  author={Shao, Songdong and Ji, Changming},
  journal={International Journal for numerical methods in fluids},
  volume={51},
  number={8},
  pages={913--936},
  year={2006},
  publisher={Wiley Online Library}
}
@article{violeau2007numerical,
  title={Numerical modelling of complex turbulent free-surface flows with the SPH method: an overview},
  author={Violeau, Damien and Issa, Reza},
  journal={International Journal for Numerical Methods in Fluids},
  volume={53},
  pages={277--304},
  year={2007}
}

@article{holm1999fluctuation,
  title={Fluctuation effects on 3D Lagrangian mean and Eulerian mean fluid motion},
  author={Holm, Darryl D},
  journal={Physica D: Nonlinear Phenomena},
  volume={133},
  number={1},
  pages={215--269},
  year={1999},
  publisher={Elsevier}
}

@article{monaghan2011turbulence,
  title={A turbulence model for Smoothed Particle Hydrodynamics},
  author={Monaghan, JJ},
  journal={European Journal of Mechanics-B/Fluids},
  volume={30},
  number={4},
  pages={360--370},
  year={2011},
  publisher={Elsevier}
}

@article{violeau2007numerical,
  title={Numerical modelling of complex turbulent free-surface flows with the SPH method: an overview},
  author={Violeau, Damien and Issa, Reza},
  journal={International Journal for Numerical Methods in Fluids},
  volume={53},
  pages={277--304},
  year={2007}
}

@article{issa2005numerical,
  title={Numerical assessment of the Smoothed Particle Hydrodynamics gridless method for incompressible flows and its extension to turbulent flows},
  author={Issa, Reza},
  year={2005}
}

@article{monaghan2002sph,
  title={SPH compressible turbulence},
  author={Monaghan, JJ},
  journal={Monthly Notices of the Royal Astronomical Society},
  volume={335},
  number={3},
  pages={843--852},
  year={2002},
  publisher={Oxford University Press}
}

%@book{gatski2013compressibility,
%  title={Compressibility, turbulence and high speed flow},
%  author={Gatski, Thomas B and Bonnet, Jean-Paul},
%  year={2013},
%  publisher={Academic Press}
%}

@misc {NASACompressibleTurbulence,
	title = {Implementing Turbulence Models into the Compressible RANS Equations},
	month = {Agu},
	url={http://turbmodels.larc.nasa.gov/implementrans.html},
	year = {2014},
	author = {Christopher Rumsey}
}

@book{cebeci2013analysis,
  title={Analysis of turbulent flows with computer programs},
  author={Cebeci, Tuncer},
  year={2013},
  publisher={Butterworth-Heinemann}
}

@article{kays1994turbulent,
  title={Turbulent Prandtl number—where are we?},
  author={Kays, William M},
  journal={Journal of Heat Transfer},
  volume={116},
  number={2},
  pages={284--295},
  year={1994},
  publisher={American Society of Mechanical Engineers}
}

%%SPH parallel
@article{goozee2003distributed,
  title={Distributed and shared memory parallelism with a smoothed particle hydrodynamics code},
  author={Gooz{\'e}e, Richard J and Jacobs, Peter A},
  journal={ANZIAM Journal},
  volume={44},
  pages={202--228},
  year={2003}
}

@article{holmes2011framework,
  title={A framework for parallel computational physics algorithms on multi-core: SPH in parallel},
  author={Holmes, David W and Williams, John R and Tilke, Peter},
  journal={Advances in Engineering Software},
  volume={42},
  number={11},
  pages={999--1008},
  year={2011},
  publisher={Elsevier}
}

@inproceedings{wenbo2014performance,
  title={Performance analysis of parallel smoothed particle hydrodynamics on multi-core CPUs},
  author={Wenbo, Chen and Yao, Yucheng and Zhang, Yang},
  booktitle={Cloud Computing and Internet of Things (CCIOT), 2014 International Conference on},
  pages={85--90},
  year={2014},
  organization={IEEE}
}

@inproceedings{adams2007adaptively,
  title={Adaptively sampled particle fluids},
  author={Adams, Bart and Pauly, Mark and Keiser, Richard and Guibas, Leonidas J},
  booktitle={ACM Transactions on Graphics (TOG)},
  volume={26},
  number={3},
  pages={48},
  year={2007},
  organization={ACM}
}

@article{dominguez2011optimization,
  title={Optimization strategies for parallel CPU and GPU implementations of a meshfree particle method},
  author={Dom{\'\i}nguez, Jose M and Crespo, Alejandro JC and G{\'o}mez-Gesteira, Moncho},
  journal={arXiv preprint arXiv:1110.3711},
  year={2011}
}

@article{ferrari2009new,
  title={A new 3D parallel SPH scheme for free surface flows},
  author={Ferrari, Angela and Dumbser, Michael and Toro, Eleuterio F and Armanini, Aronne},
  journal={Computers \& Fluids},
  volume={38},
  number={6},
  pages={1203--1217},
  year={2009},
  publisher={Elsevier}
}

@article{crespo2015dualsphysics,
  title={DualSPHysics: Open-source parallel CFD solver based on Smoothed Particle Hydrodynamics (SPH)},
  author={Crespo, AJC and Dom{\'\i}nguez, JM and Rogers, BD and G{\'o}mez-Gesteira, M and Longshaw, S and Canelas, R and Vacondio, R and Barreiro, A and Garc{\'\i}a-Feal, O},
  journal={Computer Physics Communications},
  volume={187},
  pages={204--216},
  year={2015},
  publisher={Elsevier}
}

%Dr. Patra's paper
@article{laszloffy2000simple,
  title={Simple data management, scheduling and solution strategies for managing the irregularities in parallel adaptive hp finite element simulations},
  author={Laszloffy, Andras and Long, Jingping and Patra, Abani K},
  journal={Parallel Computing},
  volume={26},
  number={13},
  pages={1765--1788},
  year={2000},
  publisher={Elsevier}
}

@article{patra2003data,
  title={Data structures and load balancing for parallel adaptive hp finite-element methods},
  author={Patra, AK and Laszloffy, A and Long, J},
  journal={Computers \& Mathematics with Applications},
  volume={46},
  number={1},
  pages={105--123},
  year={2003},
  publisher={Elsevier}
}

@article{patra2005parallel,
  title={Parallel adaptive numerical simulation of dry avalanches over natural terrain},
  author={Patra, Abani K and Bauer, AC and Nichita, CC and Pitman, E Bruce and Sheridan, MF and Bursik, M and Rupp, B and Webber, A and Stinton, AJ and Namikawa, LM and others},
  journal={Journal of Volcanology and Geothermal Research},
  volume={139},
  number={1},
  pages={1--21},
  year={2005},
  publisher={Elsevier}
}

@article{pitman2003computing,
  title={Computing granular avalanches and landslides},
  author={Pitman, E Bruce and Nichita, C Camil and Patra, Abani and Bauer, Andy and Sheridan, Michael and Bursik, Marcus},
  journal={Physics of Fluids (1994-present)},
  volume={15},
  number={12},
  pages={3638--3646},
  year={2003},
  publisher={AIP Publishing}
}

@article{patra1995problem,
  title={Problem decomposition for adaptive hp finite element methods},
  author={Patra, Abani and Oden, J Tinsley},
  journal={Computing Systems in Engineering},
  volume={6},
  number={2},
  pages={97--109},
  year={1995},
  publisher={Elsevier}
}

@inproceedings{patra1999efficient,
  title={Efficient mesh partitioning for adaptive hp finite element meshes},
  author={Patra, AK and Kim, DW},
  booktitle={In International Conference on Domain Decomposition Methods},
  year={1999},
  organization={Citeseer}
}

%SFC 
@book{sagan2012space,
  title={Space-filling curves},
  author={Sagan, Hans},
  year={2012},
  publisher={Springer Science \& Business Media}
}

%domain partition
@incollection{biswas1999experiments,
  title={Experiments with repartitioning and load balancing adaptive meshes},
  author={Biswas, Rupak and Oliker, Leonid},
  booktitle={Grid Generation and Adaptive Algorithms},
  pages={89--111},
  year={1999},
  publisher={Springer}
}